\Title

### 2022-09-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

Let us first review some existing spiking neural network models. A first motivation for these models was to understand biological observations at the different scales of the central nervous system. For example, the Hodgkin-Huxley model describes the dynamics of the membrane potential of a single nerve cell and specifically explains the role of the different voltage-dependent ion channels. It also explains how the model can produce a spike. Larger scale models introduce an simpler dynamics and an explicit spiking mechanism, for example in the Integrate-and-Fire model. It predicts the emergent states of a recurrently connected neural network~\citep{brunel_phase_2000}. Several models have increased the scale of the system to include more neurons, such as The Virtual Brain (TVB)~\citep{sanz_leon_virtual_2013}, which mixes different scales and includes spiking neural networks with (analog) neural mass models. However, these methods are mainly descriptive with applications in neuroscience but few spiking neurall models have had an application in computer vision. Another class of approaches have designed normative models of spiking neural networks that aim to have applications in machine learning. One of these is the SpikeNet algorithm which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}. Another type of spiking neural network using precise spike timing attempts to determine the structure of the network in order to minimize a cost function. This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extanded in novel gradient-based methods. The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}. However, the performance of SNNs is still lagging that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.

A new generation of algorithms has emerged with the advent of so-called silicon retinas, which mimics the output of the retina by representing visual information using spikes. %A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance. 
The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations. It also offer the possiblity to use the precise timing of events. In the HOTS model for instance, the event stream induces spatio-temporal relationships between events which are represented by building ``time surfaces'', 2D images computed using the time difference to the last recorded events~\citep{lagorce_hots_2017}. In a recent study, we have introduced a spiking neural network as a classification model applied to event-based streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model. By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is, in an online fashion. We have demonstrated on several benchmark datasets that it provides with efficient computations resulting in ultra-fast categorization. Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware. A series of similar algorithms are applied to real-life applications such as optical flow~\citep{benosman_asynchronous_2012}, depth estimation~\citep{dardelet_event-by-event_2021}, or gesture recognition~\citep{maro_event-based_2020}. Recently, such an algorithm was used in the CASPR mission on the International Space Station, illustrating, given the stringent selection processes used by the space sector, that such algorithms have reached maturity for industry.

@@

Let us now formally define the heterogeneous model, as well as the task and how we will learn to solve it. First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neural network (SNN) with heterogeneous delays. Our SNN model will be tested on a computationally relevant task of motion detection in an event-based sensory input. By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory. Solving this task is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina. The task is thus to guess the motion as accurately as possible (see \seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays. To this end, we will extend the SNN model to be efficient for the resolution of the task by including convolutional kernels. We will finally describe its implementation on conventional computers. 

>>>>

Let us first review some existing spiking neural network models. A first motivation for these models was to understand biological observations at the different scales of the central nervous system. For example, the Hodgkin-Huxley model describes the dynamics of the membrane potential of a single nerve cell and specifically explains the role of the different voltage-dependent ion channels. It also explains how the model can produce a spike. Larger scale models introduce an simpler dynamics and an explicit spiking mechanism, for example in the Integrate-and-Fire model. It predicts the emergent states of a recurrently connected neural network~\citep{brunel_phase_2000}. Several models have increased the scale of the system to include more neurons, such as The Virtual Brain (TVB)~\citep{sanz_leon_virtual_2013}, which mixes different scales and includes spiking neural networks with (analog) neural mass models. However, these methods are mainly descriptive with applications in neuroscience but few spiking neural models have had an application in computer vision. Another class of approaches have designed normative models of spiking neural networks that aim to have applications in machine learning. One of these is the SpikeNet algorithm which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}. Another type of spiking neural network using precise spike timing attempts to determine the structure of the network in order to minimize a cost function. This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extanded in novel gradient-based methods. The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}. However, the performance of SNNs is still lagging that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.

A new generation of algorithms has emerged with the advent of so-called silicon retinas, which mimics the output of the retina by representing visual information using spikes. %A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance. 
The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations. It also allows using the precise timing of events. In the HOTS model for instance, the event stream induces spatio-temporal relationships between events which are represented by building ``time surfaces'', 2D images computed using the time difference to the last recorded events~\citep{lagorce_hots_2017}. In a recent study, we have introduced a SNN as a classification model applied to event-based streams using Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model. By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is, in an online fashion. We have demonstrated on several benchmark datasets that it provides with efficient computations, resulting in ultra-fast categorization. Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware. A serie of similar algorithms are applied to real-life applications such as optical flow~\citep{benosman_asynchronous_2012}, depth estimation~\citep{dardelet_event-by-event_2021}, or gesture recognition~\citep{maro_event-based_2020}. Recently, such an algorithm was used in the CASPR mission on the International Space Station, illustrating, given the stringent selection processes used by the space sector, that such algorithms have reached maturity for industry.

@@

Let us now formally define the heterogeneous model, as well as the task and how we will learn to solve it. First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays. Our SNN model will be tested on a computationally relevant task of motion detection in an event-based sensory input. By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory. Solving this task is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina. The task is thus to guess the motion as accurately as possible (\seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays. To this end, we will extend the SNN model to be efficient for the resolution of the task by including convolutional kernels. We will finally describe its implementation on conventional computers. 
